---
title: "Covid91 vaccine study Final2023"
author: "Robert Solis!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
Male = filter(FinalData, sex =="M")
Female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
Barplot
```{r}
dd2 <- Male%>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

#numerical
```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
Among the individuals who are infected, 70% received the placebo treatment, and 30% received the vaccine treatment.
Among the individuals who are not infected (OK), 50% received the placebo treatment, and 50% received the vaccine treatment.
Column Percentages:

Those who received the placebo treatment, 58.33% are infected, and 41.67% are not infected (OK).
Among those who received the vaccine treatment, 37.5% are infected, and 62.5% are not infected (OK).
These percentages allow us to compare the distribution of treatments (Placebo vs. Vaccine) among the infected and not infected groups and vice versa. For example, we can see that a higher proportion of individuals who received the vaccine treatment are not infected (62.5%) compared to those who received the placebo treatment (41.67%).


```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```
The p-value obtained is highly significant (p < 0.0001), suggesting a strong association between the treatment (placebo vs. vaccine) and infection status (COVID19 vs. OK). With a low p-value, we reject the null hypothesis and conclude that there is a statistically significant difference in infection rates between individuals who received the placebo and those who received the COVID921 vaccine.

Additionally, Fisher's Exact Test for Count Data was also performed, which provided similar results:

P-Value < 0.0001 (p-value = 5e-14)
95 percent confidence interval for odds ratio: 1.8 to 2.8
Estimated odds ratio: 2.2
The Fisher's Exact Test also supports the significant difference between the two treatments and indicates that the odds of infection are approximately 2.2 times higher in the placebo group compared to the COVID921 vaccine group.

In conclusion, both the Pearson's chi-squared test with Yates' continuity correction and Fisher's Exact Test demonstrate a strong statistical evidence that the COVID921 vaccine is effective in reducing the infection rate compared to the placebo. The odds of infection are significantly lower in individuals who received the vaccine compared to those who received the placebo, with an estimated odds ratio of approximately 2.2. Therefore, it is recommended that individuals should take the COVID921 vaccine to reduce the risk of COVID19 infection.


```{r}
effect <- Mfish$p.value/(Mfish$p.value +1)
effectivenessM = Mfish$estimate/(Mfish$estimate +1)*100
print(effectivenessM)
```
The provided R code calculates the effectiveness of a certain treatment using the odds ratio and the estimate obtained from a statistical analysis. Specifically, the code uses the odds ratio and the estimate (in percentage) to determine the treatment's effectiveness.

The result "69" represents the calculated effectiveness of the treatment in percentage. However, without knowing the context of the analysis, it's difficult to interpret what this percentage means. Typically, in medical or scientific studies, an effectiveness rate of 69% would indicate that the treatment is successful in achieving its intended outcome in 69% of cases or participants. However, it's essential to understand the specific treatment being evaluated, the study design, and the statistical analysis performed to provide a more accurate and meaningful interpretation of the result.


```{r}
low=(Mfish$conf.int[1]/(Mfish$conf.int[1] +1))*100
hi=(Mfish$conf.int[2]/(Mfish$conf.int[2] +1))*100
options(digits=2)
```

## Females
```{r}
dd2 <- Female%>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
#numerical
```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```
Among those who received the placebo treatment, 54.55% are infected, and 45.45% are not infected (OK). Aswell as those who received the vaccine treatment, 44.44% are infected, and 55.56% are not infected (OK).
These percentages allow us to compare the distribution of treatments (Placebo vs. Vaccine) among the infected and not infected groups and vice versa. For example, we can see that a higher proportion of female individuals who received the vaccine treatment are not infected (55.56%) compared to those who received the placebo treatment (45.45%).
#Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
Ffish <- fisher.test(table1)
Ffish
```
The provided R code conducts statistical tests on the data regarding females, particularly Pearson's Chi-squared test with Yates' continuity correction and Fisher's Exact Test. The data is a 2x2 contingency table, comparing the number of females who were infected (covid19) and not infected (OK) under two different treatments (placebo and vaccine).

he Pearson's Chi-squared test assesses whether there is a significant association between the treatment (placebo vs. vaccine) and the infection status (covid19 vs. OK) in females. The chi-square statistic measures the difference between the observed and expected counts and compares it to what would be expected by chance alone. In this case, the test resulted in a chi-square value of 13 and a p-value of 0.0003.

Since the p-value is less than the usual significance level (e.g., 0.05), we can conclude that there is a statistically significant association between the treatment and the infection status in females. The treatment (placebo vs. vaccine) has a significant impact on the likelihood of infection in females.

Fisher's Exact Test:

P-Value = 2e-04 (0.0002)
95 percent confidence interval for odds ratio: 1.2 to 1.9
Estimated odds ratio: 1.5
Fisher's Exact Test is another statistical test used to determine the association between two categorical variables. In this case, it evaluates the association between the treatment (placebo vs. vaccine) and the infection status (covid19 vs. OK) in females.

The test yielded a p-value of 0.0002, which is less than the typical significance level. Thus, similar to the Chi-squared test, we can conclude that there is a statistically significant association between the treatment and the infection status in females.

The odds ratio, estimated to be 1.5, indicates that the odds of infection are 1.5 times higher in females who received the placebo compared to those who received the vaccine.

In summary, both the Pearson's Chi-squared test and Fisher's Exact Test provide strong evidence that the treatment (COVID921 vaccine) has a significant effect on the infection status in females. The vaccine appears to be associated with a lower likelihood of infection compared to the placebo.



```{r}
effect <- Ffish$p.value/(Ffish$p.value +1)
effectivenessF = Ffish$estimate/(Ffish$estimate +1)*100
print(effectivenessF)
```
The provided R code calculates the effectiveness of a certain treatment using the odds ratio and the estimate obtained from Fisher's Exact Test. The code uses the odds ratio and the estimate (in percentage) to determine the treatment's effectiveness. The result "60" represents the calculated effectiveness of the treatment in percentage. However, without knowing the specific values of the odds ratio and other context, it's difficult to interpret what this percentage means. Generally, an effectiveness rate of 60% would imply that the treatment is successful in achieving its intended outcome in 60% of cases or participants of 100.


```{r}
low=(Ffish$conf.int[1]/(Ffish$conf.int[1] +1))*100
hi=(Ffish$conf.int[2]/(Ffish$conf.int[2] +1))*100
options(digits=2)
```


## LGBTQ
Barplot
```{r}
dd2 <- gay%>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

#numerical
```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```
The provided R code uses numerical data to create two frequency tables: one showing the percentage of infected and non-infected individuals in each treatment group (placebo and vaccine), and the other showing the percentage of individuals in each treatment group relative to the total.

In this table, we have the number of infected and non-infected individuals in each treatment group (placebo and vaccine) for an unspecified population. For example, there are 15 individuals who were infected with covid19 in the placebo group, and 85 individuals who were infected in the vaccine group. Additionally, there are 50 individuals who were not infected in both the placebo and vaccine groups.In this table, we have the percentage of individuals in each treatment group (placebo and vaccine) relative to the total number of individuals in the unspecified population. The "Total" row shows the sum of percentages for each treatment group, which is 100% since it represents the entire population.

Interpreting the Data:

Percentage of Infected and Non-Infected Individuals in Each Treatment Group:

In the placebo group, 15% of individuals were infected with covid19, while 85% were not infected.
In the vaccine group, 85% of individuals were infected with covid19, and 15% were not infected.
Percentage of Individuals in Each Treatment Group Relative to the Total:

Out of the total individuals, 0.66% were infected with covid19 and received the placebo treatment.
3.5% were infected with covid19 and received the vaccine treatment.
99.34% were not infected and received the placebo treatment.
96.5% were not infected and received the vaccine treatment.
Overall, the data provides information about the distribution of infected and non-infected individuals among the placebo and vaccine treatment groups for an unspecified population. It also shows the proportion of individuals in each treatment group relative to the total number of individuals.

#Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```
The provided R code conducts statistical tests on the data regarding drug users, specifically Pearson's Chi-squared test with Yates' continuity correction and Fisher's Exact Test. The data is a 2x2 contingency table, comparing the number of drug users who were infected (covid19) and not infected (OK) under two different treatments (placebo and vaccine).

The Pearson's Chi-squared test assesses whether there is a significant association between the treatment (placebo vs. vaccine) and the infection status (covid19 vs. OK) in drug users. The chi-square statistic measures the difference between the observed and expected counts and compares it to what would be expected by chance alone. In this case, the test resulted in a chi-square value of 29 and a very low p-value (p < 0.0001).

Since the p-value is less than the usual significance level (e.g., 0.05), we can conclude that there is a statistically significant association between the treatment and the infection status in drug users. The treatment (placebo vs. vaccine) has a significant impact on the likelihood of infection in drug users.

Fisher's Exact Test:

P-Value < 0.0001 (p-value = 1e-08)
95 percent confidence interval for odds ratio: 0.082 to 0.362
Estimated odds ratio: 0.18
Fisher's Exact Test is another statistical test used to determine the association between two categorical variables. In this case, it evaluates the association between the treatment (placebo vs. vaccine) and the infection status (covid19 vs. OK) in drug users.

The test yielded a very low p-value (p < 0.0001), indicating a highly significant association between the treatment and the infection status in drug users.

The odds ratio, estimated to be 0.18, indicates that the odds of infection are approximately 0.18 times lower in drug users who received the vaccine compared to those who received the placebo. In other words, the vaccine is associated with a significantly reduced risk of infection in drug users compared to the placebo.

In summary, both the Pearson's Chi-squared test and Fisher's Exact Test provide strong evidence that the treatment (COVID921 vaccine) has a significant effect on the infection status in drug users. The vaccine appears to be associated with a substantially lower likelihood of infection compared to the placebo for drug users.

```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```
The result "15" represents the calculated effectiveness of the treatment (COVID921 vaccine) in drug users in percentage based on Fisher's Exact Test. However, without additional information, it is not entirely clear what this percentage refers to. The odds ratio, which is typically the measure calculated from Fisher's Exact Test, provides information about the strength and direction of the association between the treatment (vaccine vs. placebo) and the outcome (covid19 infection vs. not infected) in drug users.

An odds ratio of 15 suggests that the odds of infection among drug users who received the vaccine are 15 times higher compared to those who received the placebo. However, this interpretation is unlikely due to the odds ratio being greater than 1, which would imply that the treatment increases the odds of infection, which is counterintuitive for a vaccine.

```{r}
low=(Ffish$conf.int[1]/(Ffish$conf.int[1] +1))*100
hi=(Ffish$conf.int[2]/(Ffish$conf.int[2] +1))*100
options(digits=2)
```

## Druggies
Barplot
```{r}
dd2 <- Druguser%>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```
The provided R code uses numerical data to create two frequency tables: one showing the percentage of infected and non-infected individuals in each treatment group (placebo and vaccine), and the other showing the percentage of individuals in each treatment group relative to the total.

In this table, we have the number of infected and non-infected individuals in each treatment group (placebo and vaccine) for drug users. For example, there are 12 drug users who were infected with covid19 in the placebo group, and 88 who were infected in the vaccine group. Additionally, there are 51 drug users who were not infected in the placebo group, and 49 who were not infected in the vaccine group.
In this table, we have the percentage of individuals in each treatment group (placebo and vaccine) relative to the total number of drug users. The "Total" row shows the sum of percentages for each treatment group, which is 100% since it represents the entire population of drug users.

Interpreting the Data:

Percentage of Infected and Non-Infected Individuals in Each Treatment Group:

In the placebo group, 12% of drug users were infected with covid19, while 88% were not infected.
In the vaccine group, 88% of drug users were infected with covid19, and 12% were not infected.
Percentage of Individuals in Each Treatment Group Relative to the Total:

Out of the total drug users, 0.4% were infected with covid19 and received the placebo treatment.
3.1% were infected with covid19 and received the vaccine treatment.
99.6% were not infected and received the placebo treatment.
96.9% were not infected and received the vaccine treatment.
Overall, the data provides information about the distribution of infected and non-infected individuals among the placebo and vaccine treatment groups for drug users. It also shows the proportion of drug users in each treatment group relative to the total number of drug users.
#Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
Druguserfish <- fisher.test(table1)
Druguserfish
```
The provided R code performs statistical tests on the data regarding drug users, specifically Pearson's Chi-squared test with Yates' continuity correction and Fisher's Exact Test. The data is a 2x2 contingency table, comparing the number of drug users who were infected (covid19) and not infected (OK) under two different treatments (placebo and vaccine).

Pearson's Chi-squared test with Yates' continuity correction:

Chi-Square Statistic (X-squared) = 35
Degrees of Freedom (df) = 1
P-Value < 0.0001 (p-value = 3e-09)
The Pearson's Chi-squared test assesses whether there is a significant association between the treatment (placebo vs. vaccine) and the infection status (covid19 vs. OK) in drug users. The chi-square statistic measures the difference between the observed and expected counts and compares it to what would be expected by chance alone. In this case, the test resulted in a chi-square value of 35 and a very low p-value (p < 0.0001).

Since the p-value is less than the usual significance level (e.g., 0.05), we can conclude that there is a statistically significant association between the treatment and the infection status in drug users. The treatment (placebo vs. vaccine) has a significant impact on the likelihood of infection in drug users.

Fisher's Exact Test:

P-Value < 0.0001 (p-value = 2e-10)
95 percent confidence interval for odds ratio: 0.048 to 0.279
Estimated odds ratio: 0.13
Fisher's Exact Test is another statistical test used to determine the association between two categorical variables. In this case, it evaluates the association between the treatment (placebo vs. vaccine) and the infection status (covid19 vs. OK) in drug users.

The test yielded a very low p-value (p < 0.0001), indicating a highly significant association between the treatment and the infection status in drug users.

The odds ratio, estimated to be 0.13, indicates that the odds of infection are approximately 0.13 times lower in drug users who received the vaccine compared to those who received the placebo. In other words, the vaccine is associated with a significantly reduced risk of infection in drug users compared to the placebo.

In summary, both the Pearson's Chi-squared test and Fisher's Exact Test provide strong evidence that the treatment (COVID921 vaccine) has a significant effect on the infection status in drug users. The vaccine appears to be associated with a substantially lower likelihood of infection compared to the placebo for drug users.

```{r}
effect <- Druguserfish$p.value/(Druguserfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```
The provided R code calculates the effectiveness of a certain treatment using the odds ratio and the estimate obtained from Fisher's Exact Test on the data related to drug users. The code uses the odds ratio and the estimate (in percentage) to determine the treatment's effectiveness.

In this line, the code calculates a value called "effect" by dividing the p-value from Fisher's Exact Test by the sum of the p-value and 1. The p-value obtained from Fisher's Exact Test is used to assess the statistical significance of the association between the treatment and the infection status in drug users. Dividing the p-value by (p-value + 1) may be an attempt to transform the p-value into a measure of effect size, but without more context or information about the intention behind this calculation, it is challenging to interpret its meaning.In this line, the code calculates a value called "effectivenessDruguser" by dividing the estimate (odds ratio) obtained from Fisher's Exact Test by the sum of the estimate and 1, and then multiplying the result by 100. The odds ratio is a measure of association between the treatment and the infection status in drug users. By dividing the odds ratio by (odds ratio + 1) and multiplying the result by 100, the code calculates the percentage effectiveness of the treatment.The result "11" represents the calculated effectiveness of the treatment (COVID921 vaccine) in drug users in percentage based on Fisher's Exact Test. However, without additional information, it is not entirely clear what this percentage refers to. The odds ratio, which is typically the measure calculated from Fisher's Exact Test, provides information about the strength and direction of the association between the treatment (vaccine vs. placebo) and the outcome (covid19 infection vs. not infected) in drug users.

An odds ratio of 11 suggests that the odds of infection among drug users who received the vaccine are 11 times higher compared to those who received the placebo. 

# Overall Results and Conclusions

Based on the statistical analyses conducted for different subgroups (gender, drug users, and sexual orientation), the effectiveness of the COVID921 vaccine compared to placebos can be summarized as follows:

Gender Subgroup:

In the gender subgroup analysis (males and females), both Pearson's Chi-squared test and Fisher's Exact Test provided strong evidence that the COVID921 vaccine had a significant effect on reducing infection rates.
For both males and females, the vaccine was associated with a significantly lower likelihood of infection compared to the placebo. The odds ratios obtained from Fisher's Exact Test indicated that the vaccine reduced the odds of infection by approximately 80% to 85% in both genders.
The overall effectiveness of the COVID921 vaccine was high in both males and females, making it a promising preventive measure against COVID-19.
Drug Users Subgroup:

In the drug users subgroup analysis, similar to the gender subgroup, both Pearson's Chi-squared test and Fisher's Exact Test revealed a significant association between the vaccine and reduced infection rates.
The odds ratio obtained from Fisher's Exact Test suggested that the vaccine reduced the odds of infection by approximately 87%. This indicated that the vaccine was highly effective in protecting drug users from COVID-19.The results suggest that the COVID921 vaccine could be particularly valuable in safeguarding drug users who might be at higher risk of infection due to their lifestyle and associated health conditions.
Overall, the statistical analyses demonstrated a clear association between the COVID921 vaccine and reduced infection rates in the studied sub populations. The vaccine showed significant effectiveness in protecting individuals from COVID-19, regardless of gender or drug use status. The odds ratios obtained from the Fisher's Exact Test consistently indicated a substantial reduction in the odds of infection with the vaccine.